Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-25 mr-review.Rmd
added MR egger

@greco_m_detecting_2015

@bowden_mendelian_2015

>>>>

In efforts to detect the presence of horizontal pleiotropy in two-sample MR studies with GWAS summary data, @greco_m_detecting_2015 developed the between-instrument heterogeneity Q test. In a meta-analysis of MR Wald estimates, one per SNP instrument, they calculate Q in Equation \@ref(eq:q).

\begin{equation}
Q = \sum_{k = 1}^K w_k(\hat\beta^{(k)}_{XY} - \mu_F)^2
(\#eq:q)
\end{equation}

In Equation \@ref(eq:q), $w_k$ is the inverse variance of the Wald estimator, and $\mu_F = \frac{\sum_{k = 1}^K w_k \hat\beta^{(k)}_{XY}}{\sum_{k = 1}^K w_k}$.

While the Q test tends to be conservative in small sample sizes, its power increases with increasing sample size and increasing degree of pleiotropy [@greco_m_detecting_2015].

@bowden_mendelian_2015 adapted Egger regression, originally developed 
to detect bias in meta-analyses, to detect bias from horizontal 
pleiotropy in two-sample MR studies with GWAS summary data 
[@egger1997bias]. The slope coefficient estimates the causal effect. 
Under the assumption that the SNP-exposure relationship is independent 
of the SNP's pleiotropic effects, Egger's test offers a valid test of 
the null causal hypothesis and consistently estimates the causal 
effect, even if all SNPs are invalid instruments [@bowden_mendelian_2015].
